\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{enumitem}
\setlist{noitemsep}
\begin{document}
\begin{center}
{\LARGE\bfseries Jenkins Interview Question Bank\\[4pt]Beginner -- Intermediate -- Advanced}
\\[12pt]
\small Last updated: July~18,~2025
\end{center}

\section*{Using this Document}
This bank contains \textbf{80 carefully curated questions} that progressively cover Jenkins concepts from ground level to enterprise-grade architecture and troubleshooting.  The PDF is intentionally question--only so you can practise answering aloud; consult the official documentation to verify your responses.

Questions are grouped into four tiers:
\begin{enumerate}[label=\arabic*.]
  \item Beginner (Q1--Q20)
  \item Intermediate (Q21--Q40)
  \item Advanced (Q41--Q60)
  \item Scenario--Based / Troubleshooting (Q61--Q80)
\end{enumerate}

\bigskip
\hrule\bigskip

\section*{Beginner (Q1--Q20)}
\begin{enumerate}[label=\textbf{\arabic*.}]
  \item What problem does Jenkins solve in modern software delivery?
  \item Define Continuous Integration (CI) in one sentence.
  \item List three core features of Jenkins that make it popular in DevOps pipelines.
  \item Differentiate between a \emph{freestyle} job and a \emph{Pipeline} job.
  \item How do you trigger a build manually from the Jenkins UI?
  \item Which default port does Jenkins run on, and how can you change it?
  \item Where does Jenkins store configuration data on disk?
  \item Name two plugins you install immediately after a fresh Jenkins setup and explain why.
  \item What is the purpose of the \texttt{\$WORKSPACE} environment variable?
  \item Describe the steps to schedule a nightly build at 2:30~AM.
  \item How do webhooks improve SCM polling for build triggers?
  \item What is a build artifact, and how does Jenkins archive it?
  \item Explain the meaning of \emph{blue} vs. \emph{red} build status colours.
  \item What is a \emph{Jenkinsfile} and where should it reside?
  \item State two ways to back up Jenkins safely.
  \item How can you view the console output of the last build?
  \item Name two authentication realms supported natively by Jenkins.
  \item Which plugin adds \textbf{pipeline visualisation} and an editor UI?
  \item In Jenkins terminology, what is an \emph{executor}?
  \item Give one advantage of running Jenkins in Docker.
\end{enumerate}

\section*{Intermediate (Q21--Q40)}
\begin{enumerate}[label=\textbf{\arabic*.},start=21]
  \item Explain the master--agent architecture and the purpose of node labels.
  \item Describe how to inject environment variables into a Declarative Pipeline.
  \item What are \emph{shared libraries} and why are they valuable?
  \item Illustrate use of the \texttt{when} directive to run a stage only on the \texttt{main} branch.
  \item How do you secure sensitive credentials inside a Jenkinsfile?
  \item Contrast Declarative vs. Scripted Pipeline syntax.
  \item Outline the steps to migrate Jenkins to a new server with minimal downtime.
  \item What does the Script Security plugin do and why is it needed?
  \item Describe multibranch Pipeline and its advantages for GitHub repositories.
  \item How can you throttle concurrent builds of the same job?
  \item Give an example of parameterising a Pipeline build.
  \item Explain the role of \texttt{stash} and \texttt{unstash} steps.
  \item How do you configure global tool installations (e.g., Maven, JDK)?
  \item What does the build status \emph{UNSTABLE} signify?
  \item Name two common static code analysis plugins for Jenkins.
  \item How can you publish JUnit test results from a Pipeline?
  \item Discuss the pros and cons of storing build artifacts in Jenkins vs. an external artifact repository.
  \item What is the purpose of Blue Ocean and is it still actively developed?
  \item Describe how to integrate Slack notifications that fire only on failed builds.
  \item How would you monitor Jenkins metrics using Prometheus?
\end{enumerate}

\section*{Advanced (Q41--Q60)}
\begin{enumerate}[label=\textbf{\arabic*.},start=41]
  \item Design a high--availability Jenkins topology for 2{,}000 daily builds.
  \item Explain JVM tuning parameters that improve controller performance.
  \item Describe the data flow when a GitHub webhook triggers a multibranch Pipeline.
  \item Compare the Job DSL plugin with \texttt{pipelineJob} definitions in shared libraries.
  \item Detail the steps to write, test, and publish a custom Jenkins plugin.
  \item How do you provision ephemeral Kubernetes agents for each stage?
  \item Explain Blue--Green deployment implemented through Jenkins Pipelines.
  \item Discuss best practices for securing a public--facing Jenkins instance.
  \item Show how to trigger a parameterised build via the REST API using \texttt{curl}.
  \item Describe zero--downtime Jenkins core and plugin upgrades.
  \item How can you analyse and reduce build queue wait times?
  \item Outline a disaster--recovery plan if the \texttt{JENKINS\_HOME} volume is corrupted.
  \item What is the role of the \texttt{parallel} step and what are its limitations?
  \item Explain how to integrate static application security testing (SAST) into a CI Pipeline.
  \item Discuss auditing and compliance logging for regulated industries.
  \item What anti--patterns should be avoided in massive Jenkinsfiles?
  \item Describe "Pipeline as Code" governance using pull requests and code owners.
  \item How would you implement canary deployments with Jenkins and Kubernetes?
  \item What are the implications of running builds on the controller, and how can you prevent it?
  \item Provide strategies for scaling artifact storage as build volume grows.
\end{enumerate}

\section*{Scenario--Based / Troubleshooting (Q61--Q80)}
\begin{enumerate}[label=\textbf{\arabic*.},start=61]
  \item A Pipeline fails during the Docker build stage with missing dependencies.  Walk through your triage steps.
  \item After upgrading a plugin, Pipelines throw \texttt{RejectedAccessException}.  How do you fix this?
  \item Jenkins UI becomes sluggish at peak hours.  Identify likely causes and remedies.
  \item A developer's Jenkinsfile is blocked by the sandbox.  Allow the method securely without approving everything.
  \item Agents running in Kubernetes pods occasionally hang and never reconnect.  How do you troubleshoot?
  \item Build logs reveal an exposed AWS secret key.  List immediate containment actions.
  \item Your Blue--Green deployment failed to switch traffic.  Describe rollback steps.
  \item Fan--out test execution across 10 agents, then aggregate reports in one stage.  Sketch the Pipeline.
  \item Compliance asks for an audit trail of manual approvals for production releases.  How can Jenkins satisfy this?
  \item A multibranch project creates jobs for hundreds of stale branches.  How do you prune automatically?
  \item Jenkins controller CPU spikes when the build queue is flushed.  Suggest architectural improvements.
  \item An "\texttt{exec timeout}" intermittently terminates Maven builds.  Outline investigation techniques.
  \item A credential is needed only at runtime inside a Docker container.  Inject it without committing to SCM.
  \item How would you replicate a severe production bug locally using the \texttt{replay} feature?
  \item Pipeline steps report \texttt{Too many open files}.  Propose both temporary and permanent fixes.
  \item The "\texttt{Build Now}" option disappeared for a particular job.  List possible root causes.
  \item Jenkins crashes with \texttt{java.lang.OutOfMemoryError: GC overhead limit exceeded}.  Response plan?
  \item A self--signed certificate causes Git checkouts to fail on agents.  How do you resolve quickly?
  \item Explain how to implement automatic retry with exponential backoff for flaky API tests.
  \item A REST call inside a Pipeline must go through an HTTP proxy on some agents but not others.  Handle this cleanly.
\end{enumerate}

\bigskip
\hrule
\bigskip
\begin{center}
\textit{End of Question Bank -- Happy Studying!}
\end{center}
\end{document}